---
title: "G7 Standard Projectile Drag Function"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

G7 drag coefficient table taken from [JBM Ballistics](http://www.jbmballistics.com/ballistics/downloads/text/mcg7.txt) who told Doug in e-mail they got it from the US Army Ballistic Research Labratory.

Most of this notebook depends on the drag table being loaded in a data frame named tbl with 2 columns: Mach and CD. Assuming there's a file in your working directory, you can load the data with:

```{r}
tbl<-read.table("G7 Drag Coefficients.txt", header = TRUE)
```

```{r}
plot(tbl)
```


```{r}
subsonicLimit <- 0.91
supersonicLimit <- 1.04
subsonic<-tbl[tbl$Mach < subsonicLimit,]
transonic<-tbl[tbl$Mach >= subsonicLimit  & tbl$Mach <= supersonicLimit,]
supersonic<-tbl[tbl$Mach > supersonicLimit,]
plot(tbl)
lines(subsonic, col="red")
lines(transonic, col="blue")
lines(supersonic, col="green")
```

# Helper Functions
```{r}
drawFit <- function(df, fit, color) {
  lines(unlist(df["Mach"]), predict(fit, df["CD"]), col=color)
}
polyFit <- function(df, order) {
  cd <- unlist(df["CD"])
  mach <- poly(unlist(df["Mach"]), order, raw=TRUE)
  lm(cd ~ mach)
}
drawPolyFit <- function(df, order, color) {
  drawFit(df, polyFit(df, order), color)
}
plotOrder <- function(order) {
  plot(tbl)
  drawPolyFit(subsonic, order, "red")
  drawPolyFit(transonic, order, "blue")
  drawPolyFit(supersonic, order, "green")
}
printEquation <- function(df, order) {
  cat("f(x)=")
  c=coef(polyFit(df, order))
  for(i in length(c):1){
    exponent = i-1;
    if (exponent > 1) {
      cat(sprintf("%fx^%g", c[i], i-1))
    } else if (exponent == 1) {
      cat(sprintf("%fx", c[i]))
    } else {
      cat(sprintf("%f", c[i]))
    }
    if(i != 1) {
      cat(" + ");
    }
  }
}
summarizeOrder <- function(order) {
  cat("Subsonic Summary\n")
  cat("----------------\n")
  print(summary(polyFit(subsonic, order)))
  cat("Transonic Summary\n")
  cat("-----------------\n")
  print(summary(polyFit(transonic, order)))
  cat("Supersonic Summary\n");
  cat("------------------\n")
  print(summary(polyFit(supersonic, order)))
  cat("Equations\n")
  cat("---------\n")
  printEquation(subsonic, order)
  cat(sprintf(", x < %f\n", subsonicLimit))
  printEquation(transonic, order)
  cat(sprintf(", %f <= x <= %f\n", subsonicLimit, supersonicLimit))
  printEquation(supersonic, order)
  cat(sprintf(", x > %f\n", supersonicLimit))
}
```

# Linear Fit
```{r}
plotOrder(1)
summarizeOrder(1)
```
# 2nd Order Polynomial Fit
```{r}
plotOrder(2)
summarizeOrder(2)
```



# 3rd Order Polynomial Fit
```{r}
plotOrder(3)
summarizeOrder(3)
```

# 1st Order Fitting Function

$$f(x)=0.013966x + 0.115561, x < 0.910000$$
$$f(x)=2.583600x + -2.228510, 0.910000 <= x <= 1.040000$$
$$f(x)=-0.062385x + 0.437386, x > 1.040000$$

# Verify 1nd Order Fitting Function Work as Expected
```{r}

# f(x)=0.013966x + 0.115561, x < 0.910000
# f(x)=2.583600x + -2.228510, 0.910000 <= x <= 1.040000
# f(x)=-0.062385x + 0.437386, x > 1.040000

subsonicX=seq(0, subsonicLimit, 0.01)
subsonicCD <- function(x) {
  0.013966*x + 0.115561
}

transonicX=seq(subsonicLimit, supersonicLimit, 0.01)
transonicCD <- function(x) {
  2.583600*x + -2.228510
}

supersonicX=seq(supersonicLimit, max(tbl["Mach"]), 0.01)
supersonicCD <- function(x) {
  -0.062385*x + 0.437386
}

plot(tbl)
lines(supersonicX,supersonicCD(supersonicX), col="red")
lines(transonicX,transonicCD(transonicX), col="blue")
lines(subsonicX,subsonicCD(subsonicX), col="green")
```


# 2nd Order Fitting Function

Using the coefficients from the 2nd order function, we get:

$$f(x)=0.057185x^2 -0.039590x + 0.123416, x < 0.910000$$
$$f(x)=-5.634286x^2 + 13.570457x -7.577560, 0.910000 <= x <= 1.040000$$
$$f(x)=0.012365x^2 -0.131514x + 0.519829, x > 1.040000$$


# Verify 2nd Order Fitting Function Work as Expected
```{r}

# f(x)=0.057185x^2 + -0.039590x + 0.123416, x < 0.910000
# f(x)=-5.634286x^2 + 13.570457x + -7.577560, 0.910000 <= x <= 1.040000
# f(x)=0.012365x^2 + -0.131514x + 0.519829, x > 1.040000

subsonicX=seq(0, subsonicLimit, 0.01)
subsonicCD <- function(x) {
  0.057185*x^2 -0.039590*x + 0.123416
}

transonicX=seq(subsonicLimit, supersonicLimit, 0.01)
transonicCD <- function(x) {
  -5.634286*x^2 + 13.570457*x + -7.577560
}

supersonicX=seq(supersonicLimit, max(tbl["Mach"]), 0.01)
supersonicCD <- function(x) {
  0.012365*x^2 + -0.131514*x + 0.519829
}

plot(tbl)
lines(supersonicX,supersonicCD(supersonicX), col="red")
lines(transonicX,transonicCD(transonicX), col="blue")
lines(subsonicX,subsonicCD(subsonicX), col="green")
```

# Mixed Fitting Function
Mixed uses linear functions for subsonic and transonic regions, and a 2nd order function for the supersonic region.

$$f(x)=0.013966x + 0.115561, x < 0.910000$$
$$f(x)=2.583600x + -2.228510, 0.910000 <= x <= 1.040000$$

$$f(x)=0.012365x^2 + -0.131514x + 0.519829, x > 1.040000$$


# Verify Mixed Fitting Function
```{r}


subsonicX=seq(0, subsonicLimit, 0.01)
subsonicCD <- function(x) {
  0.013966*x + 0.115561
}

transonicX=seq(subsonicLimit, supersonicLimit, 0.01)
transonicCD <- function(x) {
  2.583600*x + -2.228510
}

supersonicX=seq(supersonicLimit, max(tbl["Mach"]), 0.01)
supersonicCD <- function(x) {
  0.012365*x^2 + -0.131514*x + 0.519829
}

plot(tbl)
lines(supersonicX,supersonicCD(supersonicX), col="red")
lines(transonicX,transonicCD(transonicX), col="blue")
lines(subsonicX,subsonicCD(subsonicX), col="green")
```